---
title: "Final Assignment"
author: "Daniel Galea, Konrad Karas and Long Tran, group 6"
date: "21 March 2018"
output: pdf_document
fontsize: 11pt
highlight: tango
---

### Exercise Galapagos

```{r}
galapagos = read.table("gala.txt", header = TRUE)
attach(galapagos)
```

**Point 1**

```{r}
summary(lm(Species ~ Area + Elevation + Nearest + Scruz + Adjacent, data = galapagos))
```


```{r}
summary(lm(Species ~ Area + Elevation + Scruz + Adjacent, data = galapagos))
```

```{r}
summary(lm(Species ~ Elevation + Scruz + Adjacent, data = galapagos))
```

```{r}
summary(lm(Species ~ Elevation + Adjacent, data = galapagos))
```


Using the step-down approach, we look at the highest p-value (which must also be greater than 0.05) and remove it, and run the model again. This is done until the remaining variables are significant. In this case we first removed the variable $nearest$ followed by $area$ and finally $scruz$.


**Point 2**


```{r}
summary(lm(sqrt(Species) ~ Area + Elevation + Nearest + Scruz + Adjacent, data = galapagos))
```

```{r}
summary(lm(sqrt(Species) ~ Area + Elevation + Scruz + Adjacent, data = galapagos))
```

```{r}
summary(lm(sqrt(Species) ~ Area + Elevation + Adjacent, data = galapagos))
```

```{r}
summary(lm(sqrt(Species) ~ Elevation + Adjacent, data = galapagos))


```

Using the step-down approach we first removed $nearest$ followed by $scruz$ and finally $area$.



**Point 3**

```{r}
model1 = lm(Species ~ Elevation + Adjacent, data = galapagos)
model2 = lm(sqrt(Species) ~ Elevation + Adjacent, data = galapagos)

par(mfrow = c(1, 2))
qqnorm(residuals(model1))
qqline(residuals(model1))
plot(fitted(model1), residuals(model1))

par(mfrow = c(1, 2))
qqnorm(residuals(model2))
qqline(residuals(model2))
plot(fitted(model2), residuals(model2))
```

Looking at the above plots we can conclude that the linear model with the response $sqrt(Species)$ is the better model as its residuals are more normal and the fitted plot is equally spread out.



**Point 4**

```{r}
round(cooks.distance(model2), 2)
plot(1:30, cooks.distance(model2))
```

```{r}
model3 = lm(sqrt(Species) ~ Elevation + Adjacent, data = galapagos[-16, ])
summary(model3)
```

Looking at the plot we can see that the island which is the most influential point is the island $Isabela$. After removing it from the model and running the summary we can conclude that the changes are insignificant.


**Point 5**

```{r}
par(mfrow=c(2,3))
for (i in 1:6) hist(galapagos[,i],main=colnames(galapagos)[i],xlab="",ylab="")
pairs(galapagos, panel=panel.smooth)
for (i in 1:6) hist(log(galapagos[,i]),main=colnames(galapagos)[i],xlab="",ylab="")
pairs(log(galapagos), panel=panel.smooth)
```


By applying this transformation the data seems to be more linear and it became more readable. This meant that we could see more correlations between the explanatory variables and Species, furthermore the data become more normal.



**Point 6**

```{r}
modlog=lm(log(Species)~log(Area)+log(Elevation)+log(Nearest)+log(Scruz+1)+log(Adjacent),data=galapagos)
modlog1 = step(modlog)
summary(modlog1)
```

Looking at the summary above we can conclude that not all variables are significant. The variable $log(Scruz + 1)$ is not significant.

**Point 7**
Species is exponential so applying log to it and area makes more sense since area is not exponential. 
**Point 8**

```{r}
plot(1:30, cooks.distance(modlog1))
```

```{r}
par(mfrow = c(1, 2))
qqnorm(residuals(modlog1))
qqline(residuals(modlog1))
plot(fitted(modlog1), residuals(modlog1))
```


In this model the residuals seem to be spread equally and it also seems to normalize the data and reduces influence points.

**Point 9**

```{r}
galapagos$logElevation = log(galapagos$Elevation)
galapagos$logArea = log(galapagos$Area)
galapagos$logScruz = log(galapagos$Scruz + 1)
modlog2 = lm(log(Species) ~ logArea + logElevation + logScruz, data = galapagos)

pairs(galapagos[, c(7, 8, 9)])
```

```{r}
round(cor(galapagos[, c(7, 8, 9)]), 2)
```

```{r}
summary(modlog2)
```

```{r}
modlog3 = lm(log(Species) ~ logArea * logElevation + logScruz, data = galapagos)
summary(modlog3)
```


Looking at the plot we can see that there is collinearity between elevation and area. However, after adding an interaction between them there was no significant change, therefore it is not useful to do so.


**Point 10**

