---
title: "Assignment 4"
author: "Daniel Galea, Konrad Karas and Long Tran, group 6"
date: "21 March 2018"
output: pdf_document
fontsize: 11pt
highlight: tango
---

```{r, echo = FALSE, message = FALSE}
library(knitr)
```

## Exercises


### Exercise 1


**Point 1**


**Point 2**


**Point 3**


**Point 4**


**Point 5**


**Point 6**


**Point 7**


**Point 8**


**Point 9**


**Point 10**


### Exercise 2


**Point 1**


**Point 2**


**Point 3**


**Point 4**


**Point 5**


**Point 6**


**Point 7**


**Point 8**


### Exercise 3


**Point 1**


**Point 2**


The location-scale family is a family of probability distributions which includes a location paramater and a scale parameter. The location paramter shifts the data, graphically this would mean that the data is shifted to the left or right along the x-axis. The shift paramter spreads out the data or compresses it depending on whether the shift is less than 1 or greater than 1. To check if a distribution is from the poisson distribution we can use the following formula: $$P(X=x) = e^\frac{1}{\lambda} * \frac{\lambda}{x!}\  ,x = 0,1,2,3,4,... $$ Then looking at the probabilities returned we can determine whether the distribution of X follows the Poisson distribution, and it is represented as: $$X \sim Po(\lambda)$$ Furthermore, the mean and variance for a Poisson distribution should both equal $\lambda$ and for very large $\lambda$ the $Poisson(\lambda)-distribution$ is approximately equal to a normal distribution. Therefore looking at these two samples and applying the previously mentioned probability function:
```{r}
```



**Point 3**


**Point 4**


**Point 5**

