---
title: "Assignment 3"
author: "Daniel Galea, Konrad Karas and Long Tran, group 6"
date: "14 March 2018"
output: pdf_document
fontsize: 11pt
highlight: tango
---

```{r, echo = FALSE}
library(knitr)
library(multcomp)
library(lme4)
```

## Exercises

### Exercise 1

Investigating our dataset (bread.txt) we can distinguish one numerical outcome (hours) and two factors that can be fixed at discrete number of levels (3 levels of environment [cold,intermediate,warm] and 2 levels of humidity [dry,wet]).

```{r}
data = read.table("bread.txt", header=TRUE)
data
```

Such a dataset is a good candidate to be tested by 2-way ANOVA method.

**Point 1**

Data randomization for 2-way ANOVA:

```{r}
N = 3; I = 3; J = 2
rbind(rep(1:I,each=N*J),rep(1:J,N*I),sample(1:(N*I*J)))
```

As a result we obtained 18 experimental units (18 slices of bread) with assigned levels of two factors.

**Point 2**

```{r, echo = FALSE}
boxplot(hours~environment,data=data)
```
```{r, echo = FALSE}
boxplot(hours~humidity,data=data)
```

Looking at boxplots, we can see that for intermediate and warm environments, bread slices decay much faster.
Similar observation corresponds to humidity: wet bread slices decay faster then dry ones.

```{r, echo = FALSE}
attach(data)
interaction.plot(environment,humidity,hours)
interaction.plot(humidity,environment,hours)
```

Interaction plots show that there is significant correlation between environment and humidity factors.
Bread slices decay faster when they are wet and are kept in intermediate or warm environment.

**Point 3**

Analysis of variance:
To test the variance we decided to use 2-way ANOVA as we have one numerical outcome (hours) and two factors with fixed levels (humidity and environment).

```{r}
data$environment = as.factor(data$environment)
data$humidity = as.factor(data$humidity)
dataaov = lm(hours~environment*humidity, data=data)
anova(dataaov)
```

Analysis of p-values:

* environment factor have a main effect (p-value = 2.461e-10 < 0.05) 
* humidity factor have a main effect (p-value = 4.316e-06 < 0.05)
* there is significat interaction between environment and humidity (p-value = 3.705e-07 < 0.05)

**Point 4**

Both factors, environment and humidity, interact with each other.
We are not able to test only one factor without second factor taken into account.
We cannot judge which factor has the greatest influence on the decay since factors interact with each other.

**Point 5**

```{r, echo = FALSE}
qqnorm(residuals(dataaov))
qqline(residuals(dataaov))
```

We can distinguish outliers BUT I HAVE NO IDEA HOW TO COMMENT THAT


### Exercise 2

**Point 1**

**Point 2**

**Point 3**

**Point 4**

**Point 5**

**Point 6**

**Point 7**

### Exercise 3

**Point 1**

**Point 2**

**Point 3**

**Point 4**

### Exercise 4

**Point 1**

**Point 2**

**Point 3**

**Point 4**

### Exercise 5

**Point 1**

**Point 2**

**Point 3**

**Point 4**

### Exercise 6

```{r, echo = FALSE}
data = read.table("airpollution.txt", header=TRUE)
```

**Point 1**

```{r}
pairs(data, panel=panel.smooth)
```

In scatter plots we can see some factors that are highly correlated with oxidant that is a main factor of our investigation. Some slopes are highly positive (oxidant~temperature) or negative (oxidant~wind) so linear regression model should be useful here.

**Point 2**

**Point 3**

**Point 4**

**Point 5**

### Exercise 7
